Data Exploration

### 2019-01-31 src/data_exploration.Rmd
subset backgrond data to get 10k in total per month

year
```{r}
inst_by_yr <- with(presence, table(year, institutioncode))
write.csv(inst_by_yr, file = "../output/bio/institutioncode_by_yr.csv")
inst_by_yr
```

month
```{r}
inst_by_mt <- with(presence, table(month, institutioncode))
write.csv(inst_by_mt, file = "../output/bio/institutioncode_by_mth.csv")
inst_by_mt
```

@@

>>>>

year
```{r observations by institute-year}
inst_by_yr <- with(presence, table(year, institutioncode))
write.csv(inst_by_yr, file = "../output/bio/institutioncode_by_yr.csv")
inst_by_yr
```

month
```{r observations by institute-month}
inst_by_mt <- with(presence, table(month, institutioncode))
write.csv(inst_by_mt, file = "../output/bio/institutioncode_by_mth.csv")
inst_by_mt
```

and just a year-month table
```{r observations by year-month}
year_by_mt <- with(presence, table(year, month))
write.csv(year_by_mt, file = "../output/bio/year_by_mth.csv")
year_by_mt
```

@@

```{r monthly vif with and without gear}

#with gear - for jan there is none because there is only one category of gear

#without gear
vif_nogear_jan <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = janpresab))
write.csv(vif_nogear_jan, "../output/bio/monthly_vifs/vif_nogear_jan.csv", row.names = TRUE)
vif_nogear_jan

#without gear
vif_nogear_feb <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = febpresab))
write.csv(vif_nogear_feb, "../output/bio/monthly_vifs/vif_nogear_feb.csv", row.names = TRUE)
vif_nogear_feb

#with gear
vif_gear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
vif_gear_mar

vif_gear_mar <- read.csv("../output/bio/monthly_vifs/vif_gear_mar.csv", header = TRUE)
vif_gear_mar$GVIF2Dfsq <- vif_gear_mar$GVIF..1..2.Df..^2
write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
vif_gear_mar

#without gear
vif_nogear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
write.csv(vif_nogear_mar, "../output/bio/monthly_vifs/vif_nogear_mar.csv", row.names = TRUE)
vif_nogear_mar

#with gear
vif_gear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
vif_gear_apr

vif_gear_apr <- read.csv("../output/bio/monthly_vifs/vif_gear_apr.csv", header = TRUE)
vif_gear_apr$GVIF2Dfsq <- vif_gear_apr$GVIF..1..2.Df..^2
write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
vif_gear_apr

#without gear
vif_nogear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
write.csv(vif_nogear_apr, "../output/bio/monthly_vifs/vif_nogear_apr.csv", row.names = TRUE)
vif_nogear_apr

#with gear
vif_gear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
write.csv(vif_gear_may, "../output/bio/monthly_vifs/vif_gear_may.csv", row.names = TRUE)
vif_gear_may

#without gear
vif_nogear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
write.csv(vif_nogear_may, "../output/bio/monthly_vifs/vif_nogear_may.csv", row.names = TRUE)
vif_nogear_may

#with gear
vif_gear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
write.csv(vif_gear_jun, "../output/bio/monthly_vifs/vif_gear_jun.csv", row.names = TRUE)
vif_gear_jun

#without gear
vif_nogear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
write.csv(vif_nogear_jun, "../output/bio/monthly_vifs/vif_nogear_jun.csv", row.names = TRUE)
vif_nogear_jun

#with gear
vif_gear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
vif_gear_jul

vif_gear_jul <- read.csv("../output/bio/monthly_vifs/vif_gear_jul.csv", header = TRUE)
vif_gear_jul$GVIF2Dfsq <- vif_gear_jul$GVIF..1..2.Df..^2
write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
vif_gear_jul

#without gear
vif_nogear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
write.csv(vif_nogear_jul, "../output/bio/monthly_vifs/vif_nogear_jul.csv", row.names = TRUE)
vif_nogear_jul

#with gear
vif_gear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
vif_gear_aug

vif_gear_aug <- read.csv("../output/bio/monthly_vifs/vif_gear_aug.csv", header = TRUE)
vif_gear_aug$GVIF2Dfsq <- vif_gear_aug$GVIF..1..2.Df..^2
write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
vif_gear_aug

#without gear
vif_nogear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
write.csv(vif_nogear_aug, "../output/bio/monthly_vifs/vif_nogear_aug.csv", row.names = TRUE)
vif_nogear_aug

#with gear
vif_gear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
vif_gear_sep

vif_gear_sep <- read.csv("../output/bio/monthly_vifs/vif_gear_sep.csv", header = TRUE)
vif_gear_sep$GVIF2Dfsq <- vif_gear_sep$GVIF..1..2.Df..^2
write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
vif_gear_sep

#without gear
vif_nogear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
write.csv(vif_nogear_sep, "../output/bio/monthly_vifs/vif_nogear_sep.csv", row.names = TRUE)
vif_nogear_sep

#with gear
vif_gear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
vif_gear_oct

vif_gear_oct <- read.csv("../output/bio/monthly_vifs/vif_gear_oct.csv", header = TRUE)
vif_gear_oct$GVIF2Dfsq <- vif_gear_oct$GVIF..1..2.Df..^2
write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
vif_gear_oct

#without gear
vif_nogear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
write.csv(vif_nogear_oct, "../output/bio/monthly_vifs/vif_nogear_oct.csv", row.names = TRUE)
vif_nogear_oct

#with gear
vif_gear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
vif_gear_nov

vif_gear_nov <- read.csv("../output/bio/monthly_vifs/vif_gear_nov.csv", header = TRUE)
vif_gear_nov$GVIF2Dfsq <- vif_gear_nov$GVIF..1..2.Df..^2
write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
vif_gear_nov

#without gear
vif_nogear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
write.csv(vif_nogear_nov, "../output/bio/monthly_vifs/vif_nogear_nov.csv", row.names = TRUE)
vif_nogear_nov

#with gear
vif_gear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
write.csv(vif_gear_dec, "../output/bio/monthly_vifs/vif_gear_dec.csv", row.names = TRUE)
vif_gear_dec

#without gear
vif_nogear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
write.csv(vif_nogear_dec, "../output/bio/monthly_vifs/vif_nogear_dec.csv", row.names = TRUE)
vif_nogear_dec
```

spearmans indicates mlp_surface, ssh_surface, salinity_surface, temp_surface, 02_surface all highly correlated every month. remove from each model and rerun
Run a VIF for each month (with and without gear)

```{r monthly vif with and without gear and highly correlated}

#with gear - for jan there is none because there is only one category of gear

#without gear
vif_nogear_jan <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = janpresab))
write.csv(vif_nogear_jan, "../output/bio/monthly_vifs/vif_nogear_jan.csv", row.names = TRUE)
vif_nogear_jan

#without gear
vif_nogear_feb <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface, data = febpresab))
write.csv(vif_nogear_feb, "../output/bio/monthly_vifs/vif_nogear_feb.csv", row.names = TRUE)
vif_nogear_feb

#with gear
vif_gear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
vif_gear_mar

vif_gear_mar <- read.csv("../output/bio/monthly_vifs/vif_gear_mar.csv", header = TRUE)
vif_gear_mar$GVIF2Dfsq <- vif_gear_mar$GVIF..1..2.Df..^2
write.csv(vif_gear_mar, "../output/bio/monthly_vifs/vif_gear_mar.csv", row.names = TRUE)
vif_gear_mar

#without gear
vif_nogear_mar <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = marpresab))
write.csv(vif_nogear_mar, "../output/bio/monthly_vifs/vif_nogear_mar.csv", row.names = TRUE)
vif_nogear_mar

#with gear
vif_gear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
vif_gear_apr

vif_gear_apr <- read.csv("../output/bio/monthly_vifs/vif_gear_apr.csv", header = TRUE)
vif_gear_apr$GVIF2Dfsq <- vif_gear_apr$GVIF..1..2.Df..^2
write.csv(vif_gear_apr, "../output/bio/monthly_vifs/vif_gear_apr.csv", row.names = TRUE)
vif_gear_apr

#without gear
vif_nogear_apr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
write.csv(vif_nogear_apr, "../output/bio/monthly_vifs/vif_nogear_apr.csv", row.names = TRUE)
vif_nogear_apr

#with gear
vif_gear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
write.csv(vif_gear_may, "../output/bio/monthly_vifs/vif_gear_may.csv", row.names = TRUE)
vif_gear_may

#without gear
vif_nogear_may <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = maypresab))
write.csv(vif_nogear_may, "../output/bio/monthly_vifs/vif_nogear_may.csv", row.names = TRUE)
vif_nogear_may

#with gear
vif_gear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
write.csv(vif_gear_jun, "../output/bio/monthly_vifs/vif_gear_jun.csv", row.names = TRUE)
vif_gear_jun

#without gear
vif_nogear_jun <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = junpresab))
write.csv(vif_nogear_jun, "../output/bio/monthly_vifs/vif_nogear_jun.csv", row.names = TRUE)
vif_nogear_jun

#with gear
vif_gear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
vif_gear_jul

vif_gear_jul <- read.csv("../output/bio/monthly_vifs/vif_gear_jul.csv", header = TRUE)
vif_gear_jul$GVIF2Dfsq <- vif_gear_jul$GVIF..1..2.Df..^2
write.csv(vif_gear_jul, "../output/bio/monthly_vifs/vif_gear_jul.csv", row.names = TRUE)
vif_gear_jul

#without gear
vif_nogear_jul <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = julpresab))
write.csv(vif_nogear_jul, "../output/bio/monthly_vifs/vif_nogear_jul.csv", row.names = TRUE)
vif_nogear_jul

#with gear
vif_gear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
vif_gear_aug

vif_gear_aug <- read.csv("../output/bio/monthly_vifs/vif_gear_aug.csv", header = TRUE)
vif_gear_aug$GVIF2Dfsq <- vif_gear_aug$GVIF..1..2.Df..^2
write.csv(vif_gear_aug, "../output/bio/monthly_vifs/vif_gear_aug.csv", row.names = TRUE)
vif_gear_aug

#without gear
vif_nogear_aug <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = augpresab))
write.csv(vif_nogear_aug, "../output/bio/monthly_vifs/vif_nogear_aug.csv", row.names = TRUE)
vif_nogear_aug

#with gear
vif_gear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
vif_gear_sep

vif_gear_sep <- read.csv("../output/bio/monthly_vifs/vif_gear_sep.csv", header = TRUE)
vif_gear_sep$GVIF2Dfsq <- vif_gear_sep$GVIF..1..2.Df..^2
write.csv(vif_gear_sep, "../output/bio/monthly_vifs/vif_gear_sep.csv", row.names = TRUE)
vif_gear_sep

#without gear
vif_nogear_sep <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = seppresab))
write.csv(vif_nogear_sep, "../output/bio/monthly_vifs/vif_nogear_sep.csv", row.names = TRUE)
vif_nogear_sep

#with gear
vif_gear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
vif_gear_oct

vif_gear_oct <- read.csv("../output/bio/monthly_vifs/vif_gear_oct.csv", header = TRUE)
vif_gear_oct$GVIF2Dfsq <- vif_gear_oct$GVIF..1..2.Df..^2
write.csv(vif_gear_oct, "../output/bio/monthly_vifs/vif_gear_oct.csv", row.names = TRUE)
vif_gear_oct

#without gear
vif_nogear_oct <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = octpresab))
write.csv(vif_nogear_oct, "../output/bio/monthly_vifs/vif_nogear_oct.csv", row.names = TRUE)
vif_nogear_oct

#with gear
vif_gear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
vif_gear_nov

vif_gear_nov <- read.csv("../output/bio/monthly_vifs/vif_gear_nov.csv", header = TRUE)
vif_gear_nov$GVIF2Dfsq <- vif_gear_nov$GVIF..1..2.Df..^2
write.csv(vif_gear_nov, "../output/bio/monthly_vifs/vif_gear_nov.csv", row.names = TRUE)
vif_gear_nov

#without gear
vif_nogear_nov <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = novpresab))
write.csv(vif_nogear_nov, "../output/bio/monthly_vifs/vif_nogear_nov.csv", row.names = TRUE)
vif_nogear_nov

#with gear
vif_gear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
write.csv(vif_gear_dec, "../output/bio/monthly_vifs/vif_gear_dec.csv", row.names = TRUE)
vif_gear_dec

#without gear
vif_nogear_dec <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = decpresab))
write.csv(vif_nogear_dec, "../output/bio/monthly_vifs/vif_nogear_dec.csv", row.names = TRUE)
vif_nogear_dec
```
